Parsing Postfix log files

### 2008-08-25 doc/logparser-introduction.tex
Changes made following suggestions from diction.

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check whether new anti-spam measures are working.  The
more adventurous will have generated statistics to show how many hits each
of their anti-spam measures has gotten in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this type of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\DNSBL{}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection; there are many ways in which the log lines may differ
between servers, even within the same organisation --- servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \PLP{}, a program which parses log files and places the
resulting data into a database for later use.  The gathered data can then
be used to optimise current anti-spam measures, provide a baseline to test
new anti-spam measures against, or to produce statistics showing how
effective those measures are.  There are numerous other uses for such data:
improving server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks which propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

@@

Section~\ref{parsing-algorithm} contains the core of the paper, describing
a naive parsing algorithm and the complications initially encountered which
shaped the full algorithm.  A flow chart and a discussion of the emergent
behaviour exhibited by the algorithm accompanies a comprehensive
explanation of the different stages of the initial algorithm.  The
framework which actions and rules fit into is documented, then the actions
taken during execution of the algorithm are described, followed by the
process of adding a new action.  The section concludes with an in-depth
description of the further complications discovered, and their solutions
which complete the parser.

>>>>

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how many hits each of
their anti-spam measures has had in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\DNSBL{}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection; there are many ways in which the log lines may differ
between servers, even within the same organisation --- servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \PLP{}, a program that parses log files and places the resulting
data into a database for later use.  The gathered data can then be used to
optimise current anti-spam measures, provide a baseline to test new
anti-spam measures against, or to produce statistics showing how effective
those measures are.  Numerous other uses are possible for such data:
improving server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks that propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

@@

Section~\ref{parsing-algorithm} contains the core of the paper, describing
a naive parsing algorithm and the complications initially encountered that
shaped the full algorithm.  A flow chart and a discussion of the emergent
behaviour exhibited by the algorithm accompanies a comprehensive
explanation of the different stages of the initial algorithm.  The
framework that actions and rules fit into is documented, then the actions
taken during execution of the algorithm are described, followed by the
process of adding a new action.  The section concludes with an in-depth
description of the further complications discovered, and their solutions
that complete the parser.

### 2008-08-25 doc/logparser-other-parsers-review.tex
Changes made following suggestions from diction.

Ten other parsers, starting with the only previously published research in
this area, have been reviewed as part of the background research for this
project.  None of the parsers reviewed perform the type of advanced parsing
and log correlation required for this project; all bar one are intended to
perform a specific parsing and reporting task, rather than to be a generic
parser, extracting data and leaving generation of reports from the data to
other programs.  Some parsers save the data extracted to a data store, but
the majority discard all data once they have finished running, making
historical analysis impossible.  The other parsers reviewed all produce a
report of varying complexity and detail, whereas the program described here
does not attempt to produce a report at all; that responsibility is
deferred to a separate program, to be developed later.  The parsing
algorithm and program developed for this project are designed to enable
much more detailed log analysis by providing a stable platform for
subsequent programs to develop upon.

@@

There only appears to be one prior paper published about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL or Berkeley DB database.  The decision to
support both \CSV{} and Berkeley DB appears to have been a limiting factor:
\CSV{} is a very simple format where a record is stored in a single line,
with fields separated by a comma or other punctuation symbol.  Problems
with \CSV{} files include the difficulty in escaping separators, providing
multiple values for a field (e.g.\ multiple recipients) and adding new
fields.  There is no facility to document the fields or the separator,
unlike \SQL{} databases where every database includes a schema naming the
fields and the type of data they store (integer, text, timestamp, etc.).

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \SQL{} databases which store arbitrary tuples; in \LMA{}'s main
table the key is an integer referred to by secondary tables, and the value
used is a \CSV{} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date and by-\IP{} tables use the
sender/recipient/date/\IP{} as the key, and a \CSV{} list of integers
referring to the main table is the value.  In effect this re-implements
\SQL{} foreign keys, but without the functionality offered by even the most
basic of \SQL{} databases (joins, ordering, searches, etc.).  It also
requires custom code to search on some combination of the above, though the
authors of \LMA{} did provide some queries: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  Berkeley DB appears to be the least useful
of the three output formats: it does not provide the functionality of a
basic \SQL{} database, not can it be used with standard Unix text
processing tools.

@@

    \item Some records have an \SMTP{} code of \textit{0}, though this is
        not a valid \SMTP{} code; presumably this is due to \LMA{} not
        parsing a line properly.  \LMA{} considers this code to be a
        rejection, so the command above must, too.

    \item In a tiny proportion of records the \SMTP{} code is
        \textit{deferred\/} (again, not a valid code, and most likely due
        to mis-parsing by \LMA{}); it is treated as a rejection by \LMA{},
        so the equivalent \texttt{awk} command must do so, too.

@@

There are a number of differences in the output from DAILY-REJECT and the
\texttt{awk} command:

@@

    \item Where there are multiple records in the \CSV{} file, DAILY-REJECT
        produces only one output line, whereas the \texttt{awk} command
        produces all matching lines.

@@

Pflogsumm produces a report designed for troubleshooting, rather than for
in-depth analysis.  It does not support saving any data, nor does it
extract more data than is required for producing the report.  Both the
parsing and reporting are difficult to extend as it is a specialised tool,
unlike the easily extensible design of \parsername{}.  It does not attempt
to correlate log lines by queueid or \pid{}, nor does it need to deal with
the complications encountered during this project.  Pflogsumm produces a
useful report, and successfully dealt with the \numberOFlogFILES{} log
files files tested with,\footnote{The results it reported were not
verified, but it did not report any errors, and has a very good reputation
amongst Postfix users.} but is not a suitable base for this project.

@@

There are many options available to include or exclude certain sections of
the report.

@@

Sawmill is a general purpose commercial product which parses 687 log file
formats (correct as of 2007/12/04) and produces reports.  Its data
extraction facilities are quite limited, though it does extract three
different sets of data for Postfix (one is beta as of 2007/12/04), but they
do not appear to be interlinked, nor does it save sufficient data for the
purposes of this project.\footnote{The data extracted by Sawmill is
described later in this review, and the data extracted by \PLP{} is
described in \sectionref{connections table} and \sectionref{results
table}.}  No attempt is made to correlate log lines or deal with the
difficulties documented in \sectionref{complications}
\sectionref{additional complications}.\footnote{If any attempt is made
there is no reference to it in the documentation available on the website.}
The source code is available in an obfuscated form only (presumably for a
fee), and the product is quite expensive, as it requires a license per
report which is to be generated; in contrast \parsername{} is free and the
code is freely available.  The web interface allows creation of dynamic
reports based on any field, but due to the associated cost the author has
not experimented with it.  Presumably, as the company charges per report
the user wishes to generate, the data store (if there is one) is
deliberately inaccessible and undocumented to prevent the user bypassing
the program and generating their own reports.

@@

Sawmill's web interface supports searching on any combination of the fields
it extracts, but the interface is neither as simple to use nor as
informative as the interface provided by Splunk.  The administrative
interface is very easy to use --- it took only five minutes to start
parsing a directory of log files.

@@

Splunk aims to index all an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved
searches\footnote{The author was unable to save searches, though that may
have been due to limitations in the free version.} can be run periodically
and the results emailed to a recipient or sent to a shell script, which
presumably can publish the results as required (though possibly without the
graphs and detailed statistics); \parsername{} provides the database and
leaves it to the user to utilise it in any way, whenever they want, with no
restrictions on usage of the data.  The interface is optimised for
interactive rather than automated queries and it does not appear to be
possible to write independent tools to utilise the Splunk database; there
are additional reports available on \url{http://www.splunkbase.com/}, but
the author was unable to find any: every category was empty, even those
which the interface claimed had reports available.  Many types of reports
are bundled with the software, though most are variations of a bar or pie
chart, with the exception of bubble and heatmap graphs.  It is very easy to
drill down through the graphs to extract a portion of the data (e.g.\
select the hour with the largest number of events, then select a particular
host, and finally a specific address), though it is not possible to search
on partial words.

The web interface is quite attractive and easy to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing it took roughly 30 minutes to add a single log file to
be indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all log files in a
particular directory, but after three hours, numerous futile attempts, and
reading all available documentation the author admitted defeat.  Using the
\CLI{} was more successful: invoking the command \newline \tab{}
\texttt{splunk find logs }\textit{log-directory\/}\newline added 40 of the
\numberOFlogFILES{} log files to the queue for indexing.  Repeated attempts
enqueued the same 40 log files, without explanation as to why the others
were excluded.\footnote{The log files appear to have only been indexed
once; presumably Splunk keeps track of the files it has indexed and
discards requests to index files for a second time.  This may or may not be
a useful feature for \parsername{}.} There did not appear to be an option
to order the log files to ensure they would be processed in the order they
were created.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.  All
searches performed using the indexed data returned reasonable results.

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available which allows
indexing of up to 5GB of data per day.  The cheapest licensed version costs
\$5000 (plus \$1000 support), and still limits the volume of data indexed
per day to 500MB\@.

In the specific case of parsing Postfix log files, Splunk extracts some
standard fields: to and from addresses, HELO hostname, date, host the log
files were collected from, and protocol.  It parses the standard syslog
fields at the beginning of the line, and extracts any \texttt{key=<value>}
pairs occurring after the standard syslog prologue; these pairs are the
fields listed above.  Searches can be based on the extracted fields, and
all the text in the line is also available for searching.  \parsername{}
extracts noticeably more data, though it does not make the full text of the
line available, and the full power of \SQL{} is available when searching,
allowing the user to search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields; most importantly it makes
no attempt to correlate log lines by queueid or \pid{}, nor to handle any
of the myriad complications discussed in this document.  Some additional
Postfix reports are supposedly available at
\url{http://www.splunkbase.com/}, but the author was unable to find them
(or any other reports).

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code restricts
it to supporting the \LCD{}.  The data it will extract from an \MTA{} log
file is limited in comparison to \parsername{}: \newline \tab{} time2,
email, email\_r, host, host\_r, method, url, code and bytesd.\newline There
does not seem to be an explanation of any of those fields in the
documentation (\parsername{} provides copious documentation).  AWStats has
no special purpose code to deal with the intricacies of Postfix log files,
in fact it operates by coercing Postfix log files into Apache\footnote{The
Apache web server is the most popular HTTP server in use for the past 10
years; more information is available at \url{http://httpd.apache.org/}.}
format log files, for analysis by AWStats' HTTP log file parser.  The
converting parser only deals with a small portion of the log lines
generated by Postfix, silently skipping those it cannot deal with, and does
not distinguish between different types of rejection; there is no easy way
to extend it to handle new log lines.  Although though it does correlate
log lines by queueid, it does not deal with any of the other complications
described in this document.  AWStats supports saving data but the format of
the saved data is not documented (as far as the author could tell).  It
also supports reading compressed log files, but that functionality was not
tested.

@@

YAALA provides a Postfix parser which extracts the following fields from
specific log lines:

@@

In summary YAALA provides a Postfix parser which only attempts to parse the
most common Postfix log lines (but fails miserably), provides reasonably
flexible report generation from the limited data extracted, but has no
facilities to extract data for use in other tools.

@@

Lire is a general purpose log parser supporting many different types of log
file.  Its Postfix parser extracts the following data from Postfix log
files: \textit{The email servers' reports will show you the number of
deliveries and the volume of email delivered by day, the domains from which
you receive or send the most emails, the relays most used, etc.\/}; notably
rejections are not mentioned or dealt with, and unlike \parsername{} there
is no facility to extend the parser.  It supports multiple output formats
for generated reports (text, \HTML{}, \PDF{} and Excel 95) but the reports
do not appear to be customisable; \parsername{} does not produce any
reports.  Lire supports saving extracted data for later report generation,
but accessing this data from another application is undocumented; given the
source code it should be possible, with sufficient time and effort, to
access the data from an external program.

@@

It also counts the number of log lines parsed and skipped.  Log lines are
correlated based on the queueid (referred to as sessionname [sic] within
Logrep), but not by \pid{}.  The parsing is error prone: empty fields are
saved when the log line does not match the \regex{}, though it appears that
they will not overwrite existing data.  Most notably rejections are
completely ignored, making it unsuitable for the purposes of this project.
It does not attempt to address any of the complications in
\sectionref{complications} \sectionref{additional complications} except for
correlating by queueid.

Logrep does not come with any documentation, though some scant
documentation is available on its website (\parsername{} provides copious
documentation).  It requires a web browser to interact with it, ensuring
that automated log processing will be difficult, whereas automated
processing is a key part of \parsernames{} design.  Sadly all the author's
attempts to use Logrep failed, as it was unable to access the log files
selected; this appears to be a bug rather than operator error.\footnote{If
it is caused by operator error, the interface needs improvement as the
(minimal) instructions were followed as closely as possible, and multiple
attempts were made.}  As parsing failed it wasn't possible to review the
reports Logrep can generate (available in \HTML{} only), nor to examine the
(undocumented) format in which it can save extracted data for subsequent
reuse.

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, nor make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the log files; those which do not discard the data
retain it in a format inaccessible to other tools.  Nearly all of the
parsers reviewed can produce a report of greater or lesser detail and
complexity, a facility not offered by \parsername{}; reporting is deferred
to a subsequent application.  The quality of the documentation offered by
the subset of parsers which provide some varies from unusable to good; none
of the parsers provide any documentation on the format of their data stores
(if any).  Less than half of the parsers were capable of parsing the
\numberOFlogFILES{} test log files, and improving or extending the parsers
would have been quite a difficult task for any of the parsers.  Table
\refwithpage{Summary of parsers' features} provides a summary of the
parsers' features.

>>>>

Ten other parsers, starting with the only previously published research in
this area, have been reviewed as part of the background research for this
project.  None of the parsers reviewed perform the kind of advanced parsing
and log correlation required for this project; all bar one are intended to
perform a specific parsing and reporting task, rather than to be a generic
parser, extracting data and leaving generation of reports from the data to
other programs.  Some parsers save the data extracted to a data store, but
the majority discard all data once they have finished running, making
historical analysis impossible.  The other parsers reviewed all produce a
report of varying complexity and detail, whereas the program described here
does not produce a report at all; that responsibility is deferred to a
separate program, to be developed later.  The parsing algorithm and program
developed for this project are designed to enable much more detailed log
analysis by providing a stable platform for subsequent programs to develop
upon.

@@

There only appears to be one prior paper published about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL or Berkeley DB database.  The decision to
support both \CSV{} and Berkeley DB appears to have been a serious
limitation: \CSV{} is a very simple format where a record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \CSV{} files include the difficulty in escaping separators,
providing multiple values for a field (e.g.\ multiple recipients) and
adding new fields.  There is no mechanism to document the fields or the
separator, unlike \SQL{} databases where every database includes a schema
naming the fields and the type of data they store (integer, text,
timestamp, etc.).

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \SQL{} databases that store arbitrary tuples; in \LMA{}'s main
table the key is an integer referred to by secondary tables, and the value
used is a \CSV{} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date and by-\IP{} tables use the
sender/recipient/date/\IP{} as the key, and a \CSV{} list of integers
referring to the main table is the value.  This effectively re-implements
\SQL{} foreign keys, but without the functionality offered by even the most
basic of \SQL{} databases (joins, ordering, searches, etc.).  It also
requires custom code to search on some combination of the above, though the
authors of \LMA{} did provide some queries: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  Berkeley DB appears to be the least useful
of the three output formats: it does not provide the functionality of a
basic \SQL{} database, not can it be used with standard Unix text
processing tools.

@@

    \item Some records have an \SMTP{} code of \textit{0}, though this is
        not a valid \SMTP{} code; presumably this is because of \LMA{} not
        parsing a line properly.  \LMA{} considers this code to be a
        rejection, so the command above must, too.

    \item In a tiny proportion of records the \SMTP{} code is
        \textit{deferred\/} (again, not a valid code, and most likely
        because of mis-parsing by \LMA{}); it is treated as a rejection by
        \LMA{}, so the equivalent \texttt{awk} command must do so, too.

@@

There are two differences in the output from DAILY-REJECT and the
\texttt{awk} command:

@@

    \item Where there are multiple records in the \CSV{} file, DAILY-REJECT
        produces only one output line, whereas the \texttt{awk} command
        above produces all matching lines.

@@

Pflogsumm produces a report designed for troubleshooting, rather than for
in-depth analysis.  It does not support saving any data or extracting more
data than is required for producing the report.  Both the parsing and
reporting are difficult to extend as it is a specialised tool, unlike the
easily extensible design of \parsername{}.  It does not correlate log lines
by queueid or \pid{}, and does not need to deal with the complications
encountered during this project.  Pflogsumm produces a useful report, and
successfully dealt with the \numberOFlogFILES{} log files tested
with,\footnote{The results it reported were not verified, but it did not
report any errors, and has a very good reputation amongst Postfix users.}
but is not a suitable base for this project.

@@

Many options are available to include or exclude certain sections of the
report.

@@

Sawmill is a general purpose commercial product that parses 687 log file
formats (correct as of 2007/12/04) and produces reports.  Its data
extraction facilities are too limited to save sufficient data for the
purposes of this project,\footnote{The data extracted by Sawmill is
described later in this review, and the data extracted by \PLP{} is
described in \sectionref{connections table} and \sectionref{results
table}.} though it does extract three different sets of data for Postfix
(one is beta as of 2007/12/04), but they do not appear to be interlinked.
No attempt is made to correlate log lines or deal with the difficulties
documented in \sectionref{complications} \sectionref{additional
complications}.\footnote{If any attempt is made there is no reference to it
in the documentation available on the website.} The source code is
available in an obfuscated form only (presumably for a fee), and the
product is quite expensive, as it requires a license per report to be
generated; in contrast \parsername{} is free and the code is freely
available.  The web interface allows creation of dynamic reports based on
any field, but because of the associated cost the author has not
experimented with it.  Presumably, as the company charges per report the
user wishes to generate, the data store (if there is one) is deliberately
inaccessible and undocumented to prevent the user bypassing the program and
generating their own reports.

@@

Sawmill's web interface supports searching on any combination of the fields
it extracts, but the interface is neither as simple to use nor as
informative as the interface provided by Splunk.  The administrative
interface is easy to use --- it took only five minutes to start parsing a
directory of log files.

@@

Splunk aims to index all an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved
searches\footnote{The author was unable to save searches, though that may
have been due to limitations in the free version.} can be run periodically
and the results emailed to a recipient or sent to a shell script, that
presumably can publish the results as required (though possibly without the
graphs and detailed statistics); \parsername{} provides the database and
leaves it to the user to utilise it in any way, whenever they want, with no
restrictions on usage of the data.  The interface is optimised for
interactive rather than automated queries and it does not appear to be
possible to write independent tools to utilise the Splunk database; there
are additional reports available on \url{http://www.splunkbase.com/}, but
the author was unable to find any: every category was empty, even those
that the interface claimed had reports available.  Many types of reports
are bundled with the software, though most are variations of a bar or pie
chart, except bubble and heatmap graphs.  It is easy to drill down through
the graphs to extract a portion of the data (e.g.\ select the hour with the
largest number of events, then select a particular host, and finally a
specific address), though it is not possible to search on partial words.

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing it took roughly 30 minutes to add a single log file to
be indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all log files in a
particular directory, but after three hours, numerous futile attempts, and
reading all available documentation the author admitted defeat.  Using the
\CLI{} was more successful: invoking the command \newline \tab{}
\texttt{splunk find logs }\textit{log-directory\/}\newline added 40 of the
\numberOFlogFILES{} log files to the queue for indexing.  Repeated attempts
enqueued the same 40 log files, without explaining why the others
were excluded.\footnote{The log files appear to have only been indexed
once; presumably Splunk keeps track of the files it has indexed and
discards requests to index files for a second time.  This may or may not be
a useful feature for \parsername{}.} There did not appear to be an option
to order the log files to ensure they would be processed in the order they
were created.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.  All
searches performed using the indexed data returned reasonable results.

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available that allows
indexing of up to 5GB of data per day.  The cheapest licensed version costs
\$5000 (plus \$1000 support), and still limits the volume of data indexed
per day to 500MB\@.

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, date, host the log files were collected
from, and protocol.  It parses the standard syslog fields at the beginning
of the line, and extracts any \texttt{key=<value>} pairs occurring after
the standard syslog prologue; these pairs are the fields listed above.
Searches can be based on the extracted fields, and all the text in the line
is also available for searching.  \parsername{} extracts noticeably more
data, though it does not make the full text of the line available, and the
full power of \SQL{} is available when searching, allowing the user to
search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields; most importantly it makes
no attempt to correlate log lines by queueid or \pid{}, or to handle any of
the myriad complications discussed in this document.  Some additional
Postfix reports are supposedly available at
\url{http://www.splunkbase.com/}, but the author was unable to find them
(or any other reports).

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code restricts
it to supporting the \LCD{}.  The data it will extract from an \MTA{} log
file is limited in comparison to \parsername{}: \newline \tab{} time2,
email, email\_r, host, host\_r, method, url, code and bytesd.\newline There
does not seem to be an explanation of any of those fields in the
documentation (\parsername{} provides copious documentation).  AWStats has
no special purpose code to deal with the intricacies of Postfix log files,
in fact it operates by coercing Postfix log files into Apache\footnote{The
Apache web server is the most popular HTTP server in use for the past 10
years; more information is available at \url{http://httpd.apache.org/}.}
format log files, for analysis by AWStats' HTTP log file parser.  The
converting parser only deals with a small portion of the log lines
generated by Postfix, silently skipping those it cannot deal with, and does
not distinguish between different types of rejection; it would be a lot of
work to extend it to handle new log lines.  Although though it does
correlate log lines by queueid, it does not deal with any of the other
complications described in this document.  AWStats supports saving data but
the format of the saved data is not documented (as far as the author could
tell).  It also supports reading compressed log files, but that
functionality was not tested.

@@

YAALA provides a Postfix parser that extracts the following fields from
specific log lines:

@@

In summary YAALA provides a Postfix parser that only attempts to parse the
most common Postfix log lines (but fails miserably), provides reasonably
flexible report generation from the limited data extracted, but has no
facilities to extract data for use in other tools.

@@

Lire is a general purpose log parser supporting many different types of log
file.  Its Postfix parser extracts the following data from Postfix log
files: \textit{The email servers' reports will show you the number of
deliveries and the volume of email delivered by day, the domains from which
you receive or send the most emails, the relays most used, etc.\/}; notably
rejections are not mentioned or dealt with, and unlike \parsername{} there
is no facility to extend the parser.  It supports multiple output formats
for generated reports (text, \HTML{}, \PDF{} and Excel 95) but the reports
do not appear to be customisable; \parsername{} does not produce any
reports.  Lire supports saving extracted data for later report generation,
but accessing this data from another application is undocumented; given the
source code it should be possible, with enough time and effort, to access
the data from an external program.

@@

It also counts the number of log lines parsed and skipped.  Log lines are
correlated based on the queueid (referred to as sessionname [sic] within
Logrep), but not by \pid{}.  The parsing is error prone: empty fields are
saved when the log line does not match the \regex{}, though it appears that
they will not overwrite existing data.  Most notably rejections are
completely ignored, making it unsuitable for the purposes of this project.
It does not try to address any of the complications in
\sectionref{complications} \sectionref{additional complications} except for
correlating by queueid.

Logrep does not come with any documentation, though some scant
documentation is available on its website (\parsername{} provides copious
documentation).  It requires a web browser to interact with it, ensuring
that automated log processing will be difficult, whereas automated
processing is a key part of \parsernames{} design.  Sadly all the author's
attempts to use Logrep failed, as it was unable to access the log files
selected; this appears to be a bug rather than operator error.\footnote{If
it is caused by operator error, the interface needs improvement as the
(minimal) instructions were followed as closely as possible, and multiple
attempts were made.}  As parsing failed it was not possible to review the
reports Logrep can generate (available in \HTML{} only), or to examine the
(undocumented) format in which it can save extracted data for subsequent
reuse.

@@

There are other programs available which perform basic Postfix log parsing
(some to a greater level of detail than others), but few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, or make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the log files; those that do not discard the data
retain it in a format inaccessible to other tools.  Nearly all of the
parsers reviewed can produce a report of greater or lesser detail and
complexity, unlike \parsername{}.  The quality of the documentation offered
by the subset of parsers that provide some varies from unusable to good;
none of the parsers provide any documentation on the format of their data
stores (if any).  Fewer than half of the parsers were capable of parsing
the \numberOFlogFILES{} test log files, and improving or extending the
parsers would have been quite a difficult task for any of the parsers.
Table \refwithpage{Summary of parsers' features} provides a summary of the
parsers' features.
